---
title: "Analysis of Extensive Watershed Plant Community Sampling (Chapter 4)"
author: "Emily J. Rollinson"
date: "Wednesday, November 12, 2014"
output: html_document
---

This file contains analyses for plant communities surveyed by EJR in summer 2013 in the Upper Hudson watershed (small stream riparian zones).  Presence/absence data.  These analyses focus on correlations with environmental characteristics.  


Import the species list
```{r}
require(RCurl)
options(RCurlOptions = list(cainfo = system.file("CurlSSL", "cacert.pem", package = "RCurl")))
raw <- getURL("https://raw.githubusercontent.com/erollinson/Dissertation/master/2013_species_presence_list.csv") #insert the  raw URL for the data file on github here
list <- read.csv(text = raw) #read in the github file
```

Cast the list into a site by species matrix, and convert the output (default of class 'cast') to a dataframe.
```{r}
require(reshape)
matrix<-cast(list, River ~ Species, value='Presence')
comm<-as.data.frame(matrix)
```


Import the environment matrix
```{r}
require(RCurl)
options(RCurlOptions = list(cainfo = system.file("CurlSSL", "cacert.pem", package = "RCurl")))
raw <- getURL("https://raw.githubusercontent.com/erollinson/Dissertation/master/2013_sites_allenvt_quant.csv") #insert the  raw URL for the data file on github here
env<- read.csv(text = raw, check.names = FALSE, row.names=1) #read in the github file
```


Reframe the matrices so that the river names are row names and not a column (this is done on import for env)
```{r}
rownames(comm)<-comm$River
comm<-comm[,2:ncol(comm)]
```

Reframe the matrices so all the columns are numeric
```{r}
comm[,1:ncol(comm)]<-sapply(comm[,1:ncol(comm)], as.numeric)
env[,1:ncol(env)]<-sapply(env[,1:ncol(env)], as.numeric)
```


*this is the one used in environmental analyses presented*
chooses preferred environmental variables and gives corresponding CCA

````{r}
require(vegan)
env5<-env[,c(19,22,23,24,40:54,70:73)]

env5$developed<-rowSums(env5[,15:18])
env5$forest<-rowSums(env5[,11:13])
env5$cultivated<-rowSums(env5[,7:8])
env5$wetlands<-rowSums(env5[,5:6])

prefenv<-env5[,c(1:4,20:22, 24:27)]

require(vegan)
prefcca<-cca(comm, prefenv)
summary(prefcca)
bioenv(comm,prefenv, upto=3, metric="mahalanobis") #mahlanobis scales to unit variance and forces orthogonal variables
plot(prefcca)
anova.cca(prefcca)

#this gives crosses for species and circles for the sites


#do i want RDA instead?
prefrda<-rda(comm, prefenv)
summary (prefrda)
plot(prefrda)
anova.cca(prefrda, model="direct")

```

test for spatial autocorrelation
```{r}
require(vegan)
require(fossil)
#comm is the community p/a matrix

geo<-read.csv("2013_sitecoordinates.csv", header=T, row.names=1)
spatdist<-earth.dist(geo) #returns distances in kilometers
summary(spatdist) #minimum distance 2 km, median 57, mean 68, maximum 245 km
specdist<-vegdist(comm, method="bray")

#breaks<-c(25,50,75,100,125,150,175,200) #it only calculates for the first 3 of these distance classes 

#this looks for spatial autocorrelation in species comp
spatcorr<-mantel.correlog(specdist, spatdist, nperm=999) #ah, you have to put species first and spatial second

summary(spatcorr)
spatcorr
plot(spatcorr)

require(ade4)
mantel.rtest(specdist, spatdist, nrepet=999)


#spatial autocorrelation in enviromental factors?
envcorr<-mantel.correlog(envdist, spatdist, nperm=999)
summary(envcorr)
envcorr
plot (envcorr)
mantel.rtest(envdist, spatdist, nrepet=999)


#partial Mantel test including enviro variables
require(ecodist)
envdist<-dist(prefenv, method="euclidean")
envspat<-pmgram(specdist, spatdist, envdist, nperm=999)
plot(envspat)


#MRM

mrm<-MRM(specdist ~ spatdist + envdist, nperm=999)
mrm
```



PCA
```{r}

require(ade4)
pcatest<-dudi.pca(int_partcomm)
pcatest2<-prcomp(int_partcomm, scale=TRUE)
plot(pcatest2)
biplot(pcatest2)


```


test for limiting similarity vs. environmental filtering
```{r}

#use the pres/abs matrix for 2013 'comm' in the randomizer
#need a traits matrix for 2013 that matches the pres/abs matrix for 2013, use the same traits as were used in the 2012 analysis

traits<-read.csv("2013_TRY_traitsmatrix.csv", check.names=FALSE, row.names=1)

prunecomm<-comm[colnames(comm) %in% colnames(traits)]
prunetraits<-traits[colnames(traits) %in% colnames(comm)]

#FD requires the traits matrix to have traits in columns and the community matrix to have species in columns; therefore, transpose 'prunetraits'

tprunetraits<-as.data.frame(t(prunetraits))

#cut this down just to my traits of interest

pref_traits<-tprunetraits[,c(10,11,15,22,23,25,29,30,31,43,49,59,60,62,64,65,73,74,75)]

#interpolate the most commonly measured variable in the pref_traits matrix so there aren't NAs in the distance matrix
#can't get apply to work for a single column, so I did two - seed mass and rooting depth

fx=function(x){
  x[is.na(x)]=median(x, na.rm=TRUE)
  x
}

ipreftraits<-data.frame(pref_traits[,c(1:12,15:19)], apply(pref_traits[,c(13:14)], 2, fx))

#invcomm2: 53 x 60, species are columns
require(picante)
require(FD)
rep=9999
nsites=53
permFEve13<-matrix(nrow=rep, ncol=nsites)
permFDiv13<-matrix(nrow=rep, ncol=nsites)
permFDis13<-matrix(nrow=rep, ncol=nsites)
permRaoQ13<-matrix(nrow=rep, ncol=nsites)
for (i in 1:rep) {
  temp<-randomizeMatrix(prunecomm, null.model="independentswap", iterations = 1000)
  try(FDtemp<-dbFD(ipreftraits, temp, stand.x=TRUE, ord="podani", corr = "cailliez"), silent=TRUE)
  FEve = FDtemp$FEve
  FDiv = FDtemp$FDiv
  FDis = FDtemp$FDis
  RaoQ = FDtemp$RaoQ
  permFEve13[i,]<-FEve
  permFDiv13[i,]<-FDiv
  permFDis13[i,]<-FDis
  permRaoQ13[i,]<-RaoQ
  }

#cancelled after seven hours, had only done 1k.  Do not ever run this on the laptop again.  
  
```







no longer used code kept here for reference

Make separate environment matrices for the different buffer sizes (100m, 1km, 10km)
```{r}
env100m<-env[,1:38]
env1km<-env[,c(1:23,39:53)]
env10km<-env[,c(1:23,54:68)]
```

Run a CCA for quantitative environmental characteristics

Keyser Kill is assigned a value of 0 for LOI; no soil sample was collected because the site substrate was too rocky to collect, so the assumption of 0 organic material seems reasonable

[Is there a way to do spatial control instead including lat/long as variables?] n.b. lat/long not included in final analysis

CCAs for environmental datasets containing landcover at different buffer areas (other variables are unchanged among these three)
```{r}
require(vegan)
cca.env100m<-(cca(comm, env100m))
cca.env100m
cca.env1km<-(cca(comm, env1km))
cca.env1km
cca.env10km<-(cca(comm, env10km))
cca.env10km
```


plot the CCA - by default, this shows species as red crosses and rivers as open circles. See http://ecology.msu.montana.edu/labdsv/R/labs/lab12/lab12.html for additional plotting information.

```{r}
plot(cca.env100m)
plot(cca.env1km)
plot(cca.env10km)
```


Permutation tests for CCA to assess the significance of any groupings

```{r}
require(vegan)
aov1<-anova.cca(cca.env100m, alpha=0.05, beta=0.01, step=100, perm.max=999)
aov2<-anova.cca(cca.env1km, alpha=0.05, beta=0.01, step=100, perm.max=999)
aov3<-anova.cca(cca.env10km, alpha=0.05, beta=0.01, step=100, perm.max=999)
aov1
aov2
aov3
```


Identify the best subset of environmental variables with maximum (rank) correlation with community dissimilarities, where in upto=n, n=the number of environmental variables to identify

```{r}
require(vegan)
bioenv(comm, env, upto=8)
```

Subset the 100m environmental data to those that are most likely of interest (remove isothermality, etc) and sum the land cover to broader categories

```{r}
envsub<-env100m[,c(1:2,6:7,16:38)]
envsub$developed<-rowSums(envsub[,24:27])
envsub$forest<-rowSums(envsub[,19:21])
envsub$cultivated<-rowSums(envsub[,15:16])
envsub$wetlands<-rowSums(envsub[,13:14])
envsub2<-envsub[,c(1:12,28:31)]
```

CCA for the subset created above

```{r}
cca.envsub2<-(cca(comm, envsub2))
plot(cca.envsub2)
ord<-cca.envsub2

plot(ord, type = "n")
text(cca.envsub2, display="species", cex=0.7)


goodness(cca.envsub2, display = "species", statistic = "explained")


summary(cca.envsub2)

anovasub2<-anova.cca(cca.envsub2, alpha=0.05, beta=0.01, step=100, perm.max=999)
anovasub2
bioenv(comm, envsub2, upto=16)

env3<-read.csv("envsub2.csv")
row.names(env3)<-env3$X
env3<-env3[,2:18]
cca3<-cca(comm, env3)
plot(cca3)
summary(cca3)

env4<-env3[,c(7,10,11,12,13,14,17)]
cca4<-cca(comm, env4)
plot(cca4)
summary(cca4)
bioenv(comm, env4, upto=5)

```



Note from meeting with jessica

add elevation remove longitude DONE
distance to a highway
distance to a major population center
human population in a zip code?
add phylogenetic diversity
check my LOI against the % organic in complexes with multiple series (figure out which soil series i was likely in)
test pH, measure %sand? DONE
soil depth in the series?  (before hitting bedrock) organic matter?   - take from the ucdavis series profiles

multi-scale niche