---
title: "Greenhouse Flooding 1"
author: "Emily J. Rollinson"
date: "Monday, August 03, 2015"
output: html_document
---

Read in the data and replace missing values with na
```{r}
data<-read.csv("greenhouse_flood_exp1.csv", check.names=FALSE, na.strings=c("",".","NA"))
```

Survival analysis
0=still alive at end of study
1=died

from coxph output, expcoef = Hazard Ratio - multiplicative effect of covariate on hazard function. <1 = reduction in hazard, >1 = increase in hazard

in coxph, can use cluster(x) to cluster by non-independence (here, family)
https://rpubs.com/daspringate/survival
```{r}
require(survival)
surv<-read.csv("greenhouse_survivaltime.csv", check.names=FALSE, na.strings=c("", ".", "NA"))

#rename control to acontrol so it is alphabetically first and the contrasts are against the control

surv$Treatment<-as.character(surv$Treatment)
surv$Treatment<-replace(surv$Treatment, surv$Treatment=="control", "acontrol")
surv$Treatment<-as.factor(surv$Treatment)

#build the response variable, clusters by family (accounts for non-independence)

#cox proportional hazards model
S<-Surv(time=surv$StartWeek, time2=surv$EndWeek, event=surv$Death)
model<-coxph(S ~ Source + Treatment + Source:Treatment, data=surv)
summary(model)

#proportional hazards is OK if not clustered by family.

#test the assumption of proportional hazards
validate<-cox.zph(model, transform="km")
validate
plot(validate)

#individual subsets violate the assumption but the global model does not
#now what?

#try stratifying?  this one actually makes it worse
strmodel<-coxph(S~strata(Source) + Treatment + cluster(Family), data=surv)
summary(strmodel)
valstr<-cox.zph(strmodel, transform="km")
valstr

#kaplan-meier, but doesn't allow interactions

kmodel<-survfit(S ~ Source + Treatment, data=surv)
summary(kmodel)

```


Plot survival functions (requires running the function at the bottom of this code for ggsurv)

```{r}
require(ggplot2)

#plots the four treatments
plotmodel<-survfit(S ~ Treatment, data=surv)
ggsurv(plotmodel)


```


Growth analyses

so i want time nested within each?
can i treat deaths as censored wrt time?

```{r}
require(lme4)
rm<-lmer(Height ~ Week * Treatment * Source + (Week | Source:Individual) + (Week * Treatment || Source), data=data)
summary(rm)
```


Subset the data for just the final week
```{r}
end<-data[data$Week=="9",]

#and then remove rows with NA
end<-end[complete.cases(end),]
```

Analysis of endpoint data

```{r}
require(lme4)
require(arm)

#use the subsetted data for this (below)
#End Height
#ht0<-glm(Height ~ 1,  family="gaussian", data=end)
#ht1<-glm(Height ~ Source,  family="gaussian", data=end)
#ht2<-glm(Height ~ Source/Species,  family="gaussian", data=end)
#ht3<-glm(Height ~ Source/Species/Treatment,  family="gaussian", data=end)

#display(ht0)
#display(ht1)
#display(ht2)
#display(ht3)
#anova(ht0,ht1,ht2,ht3, test="Chisq")
#summary(ht3)
#test<-lmer(Height ~ Source * Treatment + (1|Block), end)
#anova(lm(Height~ Family/Source * Treatment, end))
#anova(lm(Leaves~ Family/Source * Treatment, end))


#rename control to acontrol so it is alphabetically first and the contrasts are against the control

end$Treatment<-as.character(end$Treatment)
end$Treatment<-replace(end$Treatment, end$Treatment=="control", "acontrol")
end$Treatment<-as.factor(end$Treatment)


#lme requires removing all rows with NA
model_lme<-lme(Height ~ Source * Treatment, random=~1|Block, data=end)
summary(model_lme)
anova(model_lme)

model_leaf<-lme(Leaves ~ Source * Treatment, random = ~1|Block, data=end)
summary(model_leaf)

```


Survival analysis http://www.statmethods.net/stats/frequencies.html
```{r}
#g test on total mortality - in each treatment, for each species? 

require(vcd)

surv<-read.csv("greenhouse_survival.csv", check.names=FALSE)
surv2<-surv[,c(1,3:5)]

mat<-xtabs(Dead~Source+Treatment, surv2)
assocstats(mat)

require(ca)
ca(mat)

#lor<-loddsratio(array,1,log=FALSE)
#summary(lor)

#mantelhaen.test(array)



```

Figures for time series
```{r}
#SUMMARY STATS
#require(bear) #bear is a deprecated package, will need to rewrite code


#Height
tsheight<-summarySE(data,measurevar="Height", groupvars=c("Species", "Treatment", "Week"), na.rm=TRUE)
tsheight$Species<-factor(tsheight$Species,levels=c("BiFr", "PoPe", "PrVu", "SoFl", "PoVi", "MaVu"))

#Leaves
tsleaves<-summarySE(data,measurevar="Leaves", groupvars=c("Species", "Treatment", "Week"), na.rm=TRUE)
tsleaves$Species<-factor(tsleaves$Species,levels=c("BiFr", "PoPe", "PrVu", "SoFl", "PoVi", "MaVu"))


#FIGURES
require(ggplot2)
cbPalette2 <- c("#999999", "#E69F00", "#56B4E9", "#009E73")

#Height
tsheight<-within(tsheight, Treatment<-factor(Treatment, levels=c("control", "flood", "bury", "fertilizer")))

ggplot(data=tsheight, aes(x=Week, y=Height, color=Treatment)) + facet_wrap(~Species, nrow=2) + geom_line(stat="identity", size=0.75) + geom_pointrange(aes(ymin=Height-se, ymax=Height+se), size=0.6) + scale_x_discrete(breaks=c(0,1,2,3,4,5,6,7,8,9), labels=c("0","1","2","3","4","5","6","7","8","9")) + scale_color_manual(values=cbPalette2) +theme_bw() +  theme(panel.grid.major=element_line(color = NA), panel.grid.minor=element_line(color = NA))

#Leaves
tsleaves<-within(tsleaves, Treatment<-factor(Treatment, levels=c("control", "flood", "bury", "fertilizer")))

ggplot(data=tsleaves, aes(x=Week, y=Leaves, color=Treatment)) + facet_wrap(~Species, nrow=2) + geom_line(stat="identity", size=0.75) + geom_pointrange(aes(ymin=Leaves-se, ymax=Leaves+se), size=0.6) + scale_x_discrete(breaks=c(0,1,2,3,4,5,6,7,8,9), labels=c("0","1","2","3","4","5","6","7","8","9")) + scale_color_manual(values=cbPalette2) +theme_bw() +  theme(panel.grid.major=element_line(color = NA), panel.grid.minor=element_line(color = NA))


#Survivorship
ggplot(data=tsleaves, aes(x=Week, y=N, color=Treatment)) + facet_wrap(~Species, nrow=2) + geom_line(stat="identity", size=0.75) + scale_x_discrete(breaks=c(0,1,2,3,4,5,6,7,8,9), labels=c("0","1","2","3","4","5","6","7","8","9")) + scale_color_manual(values=cbPalette2) +theme_bw() +  theme(panel.grid.major=element_line(color = NA), panel.grid.minor=element_line(color = NA))+ scale_y_discrete(breaks=c(0,1,2,3,4,5,6,7,8,9,10,11,12), labels=c("0","1","2","3","4","5","6","7","8","9", "10", "11", "12"))


#worksbutismessy
#ggplot(data=tsheight, aes(x=Week, y=Height, color=Treatment)) + facet_wrap(~Species, nrow=2) + geom_line(stat="identity") +geom_point(stat="identity") + geom_errorbar(aes(ymin=Height-se, ymax=Height+se)) + scale_x_discrete(limits=c(0,9), breaks=c(0,1,2,3,4,5,6,7,8,9), labels=c("0", "1","2","3","4","5", "6","7","8","9"))

```


Figures for the final week
```{r}
#require(bear)
#bear package no longer works so summarySE does not work; will need to use ddply

#example
#cleafC<-ddply(alltraits2, c("Origin"), summarise,
            # N = sum(!is.na(leafC)),
            # mean = mean(leafC, na.rm=TRUE),
           #  sd = sd(leafC, na.rm=TRUE),
           #  se = sd/sqrt(N))

require(plyr)
#SUMMARY STATS


#Height
#height<-summarySE(end,measurevar="Height", groupvars=c("Species", "Treatment"), na.rm=TRUE)
height<-ddply(end, c("Species", "Treatment"), summarise,
              N = sum(!is.na(Height)),
              mean = mean(Height, na.rm=TRUE),
              sd = sd(Height, na.rm=TRUE), 
              se = sd/sqrt(N))


#Stupid juggling to get the rows in non-alphabetical order
height$Species<-as.character(height$Species)
height$Species<-factor(height$Species,levels=c("BiFr", "PoPe", "PrVu", "SoFl", "PoVi", "MaVu"))
#replace NA SEs with 0 so ggplot doesn't fail (sample size of 1)
height[is.na(height)]<-0
levels(height$Species)<-c("Bidens frondosa", "Polygonum pennsylvanicum", "Prunella vulgaris", "Solidago flexicaulis", "Polygonum virginianum", "Marrubium vulgare")
levels(height$Treatment)<-c("control", "submerge", "bury", "nutrient")

#Leaves
#leaves<-summarySE(end, measurevar="Leaves", groupvars=c("Species", "Treatment"), na.rm=TRUE)
#leaves$Species<-as.character(leaves$Species)


leaves<-ddply(end, c("Species", "Treatment"), summarise,
                     N=sum(!is.na(Leaves)),
                     mean = mean(Leaves, na.rm=TRUE),
                     sd = sd(Leaves, na.rm=TRUE),
                     se=sd/sqrt(N))


leaves$Species<-factor(leaves$Species, levels=c("BiFr", "PoPe", "PrVu", "SoFl", "PoVi", "MaVu"))
leaves[is.na(leaves)]<-0
levels(leaves$Species)<-c("Bidens frondosa", "Polygonum pennsylvanicum", "Prunella vulgaris", "Solidago flexicaulis", "Polygonum virginianum", "Marrubium vulgare")
levels(leaves$Treatment)<-c("bury", "control", "nutrient", "submerge")


#FIGURES
require(ggplot2)
require(wesanderson)
palette <-wes_palette("Darjeeling")
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

#Height

height<-within(height, Treatment<-factor(Treatment, levels=c("control", "submerge", "bury", "nutrient")))

ggplot(height, aes(x=factor(Treatment), y=mean, fill=factor(Treatment))) + geom_point(stat="identity", position="dodge", size=4, aes(order=desc(Treatment))) +  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), colour="black", width=.4) + xlab("Species") + ylab("Height (cm)") + theme_bw() + theme(panel.grid.major=element_line(color = NA), panel.grid.minor=element_line(color = NA)) + facet_wrap(~Species, nrow=2)

#Leaves
leaves<-within(leaves, Treatment<-factor(Treatment, levels=c("control", "submerge", "bury", "nutrient")))

ggplot(leaves, aes(x=factor(Treatment), y=mean, fill=factor(Treatment))) + geom_point(stat="identity", position="dodge", size=4, aes(order=desc(Treatment))) +geom_errorbar(aes(ymin=mean-se, ymax=mean+se), colour="black", width=.4) + xlab("Species") + ylab("Number of leaves") + theme_bw() + theme(panel.grid.major=element_line(color = NA), panel.grid.minor=element_line(color = NA)) +facet_wrap(~Species, nrow=2)

#OLDFIGURES
#this works except the error bars
#ggplot(height[!is.na(height$Height),], aes(x=factor(Species), y=factor(Height) ,fill=factor(Treatment))) + geom_bar(stat="identity", position="dodge", aes(order=desc(Treatment))) + scale_fill_manual(values=cbPalette) +scale_color_manual(values=cbPalette) +  geom_errorbar(aes(ymin=Height-se, ymax=Height+se), colour="black", width=.2, position=position_dodge(width=0.9)) + xlab("Species") + ylab("Height (cm)") + theme_bw() + theme(panel.grid.major=element_line(color = NA), panel.grid.minor=element_line(color = NA)) 
#trying to fix
#ggplot(height, aes(x=factor(Species), y=factor(Height) ,fill=factor(Treatment))) + geom_bar(stat="identity", position="dodge", aes(order=desc(Treatment))) + scale_fill_manual(values=cbPalette) +scale_color_manual(values=cbPalette) +  geom_errorbar(aes(ymin=Height-se, ymax=Height+se), colour="black", width=.2, position=position_dodge(width=0.9)) + xlab("Species") + ylab("Height (cm)") + theme_bw() + theme(panel.grid.major=element_line(color = NA), panel.grid.minor=element_line(color = NA)) + scale_y_discrete(limits=c(0,45), breaks=c(0,5,10,15,20,25,30,35,40,45), labels=c("0", "5","10","15","20","25", "30","35","40","45"))
#this works
#ggplot(height[!is.na(height$Height),], aes(x=factor(Species), y=Height, fill=factor(Treatment))) + geom_bar(stat="identity", position="dodge", aes(order=desc(Treatment))) + scale_fill_manual(values=cbPalette) +scale_color_manual(values=cbPalette) +  geom_errorbar(aes(ymin=Height-se, ymax=Height+se), colour="black", width=.2, position=position_dodge(width=0.9)) + xlab("Species") + ylab("Height (cm)") + theme_bw() + theme(panel.grid.major=element_line(color = NA), panel.grid.minor=element_line(color = NA)) 
```



Function for suvival curves in ggplot by Edwin Thoen
http://www.r-statistics.com/2013/07/creating-good-looking-survival-curves-the-ggsurv-function/

```{r}
ggsurv <- function(s, CI = 'def', plot.cens = T, surv.col = 'gg.def',
                   cens.col = 'red', lty.est = 1, lty.ci = 2,
                   cens.shape = 3, back.white = F, xlab = 'Time',
                   ylab = 'Survival', main = ''){
 
  library(ggplot2)
  strata <- ifelse(is.null(s$strata) ==T, 1, length(s$strata))
  stopifnot(length(surv.col) == 1 | length(surv.col) == strata)
  stopifnot(length(lty.est) == 1 | length(lty.est) == strata)
 
  ggsurv.s <- function(s, CI = 'def', plot.cens = T, surv.col = 'gg.def',
                       cens.col = 'red', lty.est = 1, lty.ci = 2,
                       cens.shape = 3, back.white = F, xlab = 'Time',
                       ylab = 'Survival', main = ''){
 
    dat <- data.frame(time = c(0, s$time),
                      surv = c(1, s$surv),
                      up = c(1, s$upper),
                      low = c(1, s$lower),
                      cens = c(0, s$n.censor))
    dat.cens <- subset(dat, cens != 0)
 
    col <- ifelse(surv.col == 'gg.def', 'black', surv.col)
 
    pl <- ggplot(dat, aes(x = time, y = surv)) +
      xlab(xlab) + ylab(ylab) + ggtitle(main) +
      geom_step(col = col, lty = lty.est)
 
    pl <- if(CI == T | CI == 'def') {
      pl + geom_step(aes(y = up), color = col, lty = lty.ci) +
        geom_step(aes(y = low), color = col, lty = lty.ci)
    } else (pl)
 
    pl <- if(plot.cens == T & length(dat.cens) > 0){
      pl + geom_point(data = dat.cens, aes(y = surv), shape = cens.shape,
                       col = cens.col)
    } else if (plot.cens == T & length(dat.cens) == 0){
      stop ('There are no censored observations')
    } else(pl)
 
    pl <- if(back.white == T) {pl + theme_bw()
    } else (pl)
    pl
  }
 
  ggsurv.m <- function(s, CI = 'def', plot.cens = T, surv.col = 'gg.def',
                       cens.col = 'red', lty.est = 1, lty.ci = 2,
                       cens.shape = 3, back.white = F, xlab = 'Time',
                       ylab = 'Survival', main = '') {
    n <- s$strata
 
    groups <- factor(unlist(strsplit(names
                                     (s$strata), '='))[seq(2, 2*strata, by = 2)])
    gr.name <-  unlist(strsplit(names(s$strata), '='))[1]
    gr.df <- vector('list', strata)
    ind <- vector('list', strata)
    n.ind <- c(0,n); n.ind <- cumsum(n.ind)
    for(i in 1:strata) ind[[i]] <- (n.ind[i]+1):n.ind[i+1]
 
    for(i in 1:strata){
      gr.df[[i]] <- data.frame(
        time = c(0, s$time[ ind[[i]] ]),
        surv = c(1, s$surv[ ind[[i]] ]),
        up = c(1, s$upper[ ind[[i]] ]),
        low = c(1, s$lower[ ind[[i]] ]),
        cens = c(0, s$n.censor[ ind[[i]] ]),
        group = rep(groups[i], n[i] + 1))
    }
 
    dat <- do.call(rbind, gr.df)
    dat.cens <- subset(dat, cens != 0)
 
    pl <- ggplot(dat, aes(x = time, y = surv, group = group)) +
      xlab(xlab) + ylab(ylab) + ggtitle(main) +
      geom_step(aes(col = group, lty = group))
 
    col <- if(length(surv.col == 1)){
      scale_colour_manual(name = gr.name, values = rep(surv.col, strata))
    } else{
      scale_colour_manual(name = gr.name, values = surv.col)
    }
 
    pl <- if(surv.col[1] != 'gg.def'){
      pl + col
    } else {pl + scale_colour_discrete(name = gr.name)}
 
    line <- if(length(lty.est) == 1){
      scale_linetype_manual(name = gr.name, values = rep(lty.est, strata))
    } else {scale_linetype_manual(name = gr.name, values = lty.est)}
 
    pl <- pl + line
 
    pl <- if(CI == T) {
      if(length(surv.col) > 1 && length(lty.est) > 1){
        stop('Either surv.col or lty.est should be of length 1 in order
             to plot 95% CI with multiple strata')
      }else if((length(surv.col) > 1 | surv.col == 'gg.def')[1]){
        pl + geom_step(aes(y = up, color = group), lty = lty.ci) +
          geom_step(aes(y = low, color = group), lty = lty.ci)
      } else{pl +  geom_step(aes(y = up, lty = group), col = surv.col) +
               geom_step(aes(y = low,lty = group), col = surv.col)}
    } else {pl}
 
 
    pl <- if(plot.cens == T & length(dat.cens) > 0){
      pl + geom_point(data = dat.cens, aes(y = surv), shape = cens.shape,
                      col = cens.col)
    } else if (plot.cens == T & length(dat.cens) == 0){
      stop ('There are no censored observations')
    } else(pl)
 
    pl <- if(back.white == T) {pl + theme_bw()
    } else (pl)
    pl
  }
  pl <- if(strata == 1) {ggsurv.s(s, CI , plot.cens, surv.col ,
                                  cens.col, lty.est, lty.ci,
                                  cens.shape, back.white, xlab,
                                  ylab, main)
  } else {ggsurv.m(s, CI, plot.cens, surv.col ,
                   cens.col, lty.est, lty.ci,
                   cens.shape, back.white, xlab,
                   ylab, main)}
  pl
}


```
