Phylogenetic diversity analyses
==============================================================================================================================

Some useful code tidbits
-------------------------------------------------------

Note that in RStudio, the data fram view only displays a certain number of columns.  For a large dataset, if you need to view the entire data frame, you can use the command below to open it in a separate window.

utils::View(data)

```{r}
utils::View(matrix)
```

Phylogenetic diversity

```{r}
require(picante)
require(phytools)
tree<-read.newick("newick_2012sp.txt")
tree2<-collapse.singles(tree)
tree3<-compute.brlen(tree2)
test<-comdist(matrix, tree)


```

<<<<<<< HEAD
phylogenetic composition native vs. introduced subset of community

```{r}
require(picante)
require(phytools)

#read in the tree for all the species found in 2013
tree<-read.newick("newick_2013_newtree.txt")
tree2<-collapse.singles(tree)
plot(tree2)
=======


phylogenetic composition native vs. introduced subset of community
>>>>>>> 53b34dc728debd2f3da8b485489140685e2131a7

#read in 2013 species
comm13<-read.csv("2013_speciespresences_by_origin.csv")
comm<-comm13[,1:3]

<<<<<<< HEAD
#read in phylo formatted list of 2013 species   #this has the wrong origins assigned in it something got mis-sorted

splist<-read.csv("2013_taxa_phyloformat_columns.csv")

#transform 2013 species list to presence/absence matrix
require(reshape)
species<-cast(comm, River ~ Species, value="Presence", fun.aggregate=sum) #sum here will provide an accurate presence/absence matrix, since each species is only listed once per river in the original list, and the sum can only be zero or one.  fun.aggregate=max can be used, but will provide 1/-INF rather than 1/0 and the -INF will have to be transformed to 0.
species<-as.data.frame(species) #because FXN 'cast' returns output in class cast
row.names(species)<-species$River #to turn the first column into the row names
species<-species[,2:281] #to remove the duplicate first column now that the row names are labelled


#plot a tree with tip labels for origin ("trait" = origin)
color.plot.phylo(tree2, splist, "Origin", "Species")  #this is throwing an error about mismatched numbers of species but they aren't actually mismatched so i'm not sure what's happening

```

convert phylo object to interspecific distance matrix

```{r}
isdm<-cophenetic(tree2) #isdm = interspecific distance matrix
```

try calculating MPD between inv and nat subsets

```{r}
fulllist<-read.csv("2013_speciespresences_by_origin_sitenamepartitioned.csv", check.names=FALSE)


#omits species with unresolved origin status entirely (n=7 species; 34 observations)
#make this list into a pres/abs matrix

require(reshape)
partcomms<-cast(fulllist, PartSite ~ Species, value="Presence", fun.aggregate=mean)
row.names(partcomms)<-partcomms$PartSite
partcomms2<-partcomms[,2:274]
partcomms2<-rapply(partcomms2, f=function(x) ifelse(is.nan(x),0,x), how="replace")
partcomms3<-as.data.frame(partcomms2, check.names=FALSE)
row.names(partcomms3)<-partcomms$PartSite
comm13<-partcomms3

write.table(comm13, "2013presabsmatrix_partbyorigin_phylonames.csv", sep=",")
#fixed the genus.species to genus_species

comm<-read.csv("2013presabsmatrix_partbyorigin_phylonames.csv")
row.names(comm)<-comm$X
comm<-comm[,2:274]

require(picante)

test<-comdist(comm, isdm)

prune.sample(comm, tree2)

#phylodiversity

agetree<-read.newick("tree13_withages.new")
agetree2<-collapse.singles(agetree)
plot(agetree2)
test2<-pd(comm, agetree2)

ageisdm<-cophenetic(agetree2) #isdm = interspecific distance matrix

test3<-comdist(comm, ageisdm)


#phylodiversity without the I/N divide

#read in 2013 species

#transform 2013 species list to presence/absence matrix
comm13<-read.csv("2013_speciespresences_by_origin.csv")
comm<-comm13[,1:3]


require(reshape)
species<-cast(comm, River ~ Species, value="Presence", fun.aggregate=sum) #sum here will provide an accurate presence/absence matrix, since each species is only listed once per river in the original list, and the sum can only be zero or one.  fun.aggregate=max can be used, but will provide 1/-INF rather than 1/0 and the -INF will have to be transformed to 0.
species<-as.data.frame(species) #because FXN 'cast' returns output in class cast
row.names(species)<-species$River #to turn the first column into the row names
species<-species[,2:274] #to remove the duplicate first column now that the row names are labelled

PD<-pd(species, agetree2)

new<-match.phylo.comm(agetree2, species)

test3<-comdist(species, ageisdm)

```

=======
```{r}
require(picante)
require(phytools)

#read in the tree for all the species found in 2013
tree<-read.newick("newick_2013_newtree.txt")
tree2<-collapse.singles(tree)
plot(tree2)

#read in 2013 species
comm13<-read.csv("2013_speciespresences_by_origin.csv")
comm<-comm13[,1:3]

#read in phylo formatted list of 2013 species   #this has the wrong origins assigned in it something got mis-sorted

splist<-read.csv("2013_taxa_phyloformat_columns.csv")

#transform 2013 species list to presence/absence matrix
require(reshape)
species<-cast(comm, River ~ Species, value="Presence", fun.aggregate=sum) #sum here will provide an accurate presence/absence matrix, since each species is only listed once per river in the original list, and the sum can only be zero or one.  fun.aggregate=max can be used, but will provide 1/-INF rather than 1/0 and the -INF will have to be transformed to 0.
species<-as.data.frame(species) #because FXN 'cast' returns output in class cast
row.names(species)<-species$River #to turn the first column into the row names
species<-species[,2:281] #to remove the duplicate first column now that the row names are labelled


#plot a tree with tip labels for origin ("trait" = origin)
color.plot.phylo(tree2, splist, "Origin", "Species")  #this is throwing an error about mismatched numbers of species but they aren't actually mismatched so i'm not sure what's happening

```


convert phylo object to interspecific distance matrix

```{r}
isdm<-cophenetic(tree2) #isdm = interspecific distance matrix
```



try calculating MPD between inv and nat subsets

```{r}
fulllist<-read.csv("2013_speciespresences_by_origin_sitenamepartitioned.csv", check.names=FALSE)


#omits species with unresolved origin status entirely (n=7 species; 34 observations)
#make this list into a pres/abs matrix

require(reshape)
partcomms<-cast(fulllist, PartSite ~ Species, value="Presence", fun.aggregate=mean)
row.names(partcomms)<-partcomms$PartSite
partcomms2<-partcomms[,2:274]
partcomms2<-rapply(partcomms2, f=function(x) ifelse(is.nan(x),0,x), how="replace")
partcomms3<-as.data.frame(partcomms2, check.names=FALSE)
row.names(partcomms3)<-partcomms$PartSite
comm13<-partcomms3

write.table(comm13, "2013presabsmatrix_partbyorigin_phylonames.csv", sep=",")
#fixed the genus.species to genus_species

comm<-read.csv("2013presabsmatrix_partbyorigin_phylonames.csv")
row.names(comm)<-comm$X
comm<-comm[,2:274]

require(picante)

test<-comdist(comm, isdm)

prune.sample(comm, tree2)





#phylodiversity

agetree<-read.newick("tree13_withages.new")
agetree2<-collapse.singles(agetree)
plot(agetree2)
test2<-pd(comm, agetree2)


ageisdm<-cophenetic(agetree2) #isdm = interspecific distance matrix

test3<-comdist(comm, ageisdm)


#phylodiversity without the I/N divide

#read in 2013 species

#transform 2013 species list to presence/absence matrix
comm13<-read.csv("2013_speciespresences_by_origin.csv")
comm<-comm13[,1:3]


require(reshape)
species<-cast(comm, River ~ Species, value="Presence", fun.aggregate=sum) #sum here will provide an accurate presence/absence matrix, since each species is only listed once per river in the original list, and the sum can only be zero or one.  fun.aggregate=max can be used, but will provide 1/-INF rather than 1/0 and the -INF will have to be transformed to 0.
species<-as.data.frame(species) #because FXN 'cast' returns output in class cast
row.names(species)<-species$River #to turn the first column into the row names
species<-species[,2:274] #to remove the duplicate first column now that the row names are labelled

PD<-pd(species, agetree2)

new<-match.phylo.comm(agetree2, species)

test3<-comdist(species, ageisdm)

```


tools for making data frames match
match.phylo.comm(tree,comm)


