A vague attempt at RLQ analysis
========================================================

Read in the data
```{r}
traits<-read.csv("2013_babytraits.csv", check.names=FALSE) #check.names=FALSE stops R from putting periods in the species name (e.g. Abies.balsamea instead of Abies balsamea; this keeps the naming consistent with the species list)
row.names(traits)<-traits$Species
traits<-traits[,2:4]

#eliminate some unwanted categories -> biennial to perennial; vine to forb

traits$"Growth Habit" <-as.character(traits$"Growth Habit")
traits$"Growth Habit"[traits$"Growth Habit" == "vine"] <- "herb"
traits$"Growth Habit"[traits$"Growth Habit" == "forb"] <- "herb"
traits$"Growth Habit"[traits$"Growth Habit" == "graminoid"] <- "herb"
traits$"Growth Habit"[traits$"Growth Habit" == "shrub"] <- "woody"
traits$"Growth Habit"[traits$"Growth Habit" == "tree"] <- "woody"
traits$"Duration"<-as.character(traits$"Duration")
traits$"Duration"[traits$"Duration" == "biennial"] <- "perennial"
traits$"Growth Habit"<-as.factor(traits$"Growth Habit")
traits$"Duration"<-as.factor(traits$"Duration")

splist<-read.csv("2013_species_presence_list_for_traits.csv")

env<-read.csv("2013_environment_matrix_pared.csv")
row.names(env)<-env$River #to turn the first column into the row names
env<-env[,2:14] #to remove the duplicate first column now that the row names are labelled

#cut environmental variables down to most interesting
env2<-env[,c(1,4,8,11:12)]

```

Transform the species list into a matrix
```{r}
require(reshape)
species<-cast(splist, River ~ Species, value="Presence", fun.aggregate=sum) #sum here will provide an accurate presence/absence matrix, since each species is only listed once per river in the original list, and the sum can only be zero or one.  fun.aggregate=max can be used, but will provide 1/-INF rather than 1/0 and the -INF will have to be transformed to 0.
species<-as.data.frame(species) #because FXN 'cast' returns output in class cast
row.names(species)<-species$River #to turn the first column into the row names
species<-species[,2:256] #to remove the duplicate first column now that the row names are labelled
```

Get all the matrices into the correct shape 
```{r}
#site x species matrix requires species in rows, which it is

#site x environment matrix requires environment in rows
tenv<-t(env2)
tenv<-as.data.frame(tenv)

#trait x species matrix requires traits in rows
ttraits<-t(traits)
ttraits<-as.data.frame(traits)

```

Perform the RLQ analysis
```{r}
require(ade4)
modelsp<-dudi.coa(species)
modelenv<-dudi.hillsmith(env2, row.w = modelsp$lw)
modeltrait<-dudi.hillsmith(traits, row.w =modelsp$cw)
rlq1<-rlq(modelenv, modelsp, modeltrait)
plot(rlq1)
rlq1
summary(rlq1)
testrlq<-randtest(rlq1)
testrlq
fourthcorner.rlq(rlq1,type="Q.axes")
fourthcorner.rlq(rlq1,type="R.axes")

nrepet = 99999
fourth<-fourthcorner(env2, species, traits, modeltype=6, , p.adjust.method.G="fdr", p.adjust.method.D="fdr", nrepet = nrepet) 
plot(fourth, alpha = 0.05, stat = "D2")
summary(fourth)
```

p.adjust.methods
# c("holm", "hochberg", "hommel", "bonferroni", "BH", "BY",
#   "fdr", "none")

Additional plotting options courtesy http://www.esapubs.org/archive/ecol/E095/002/suppl-1.pdf
```{r}
par(mfrow = c(1,1))#asking R to plot 3 side by side plots
s.arrow(rlq1$c1, addaxes=TRUE, boxes=FALSE) # c1 = arrows for traits
s.arrow(rlq1$l1, addaxes=TRUE, add.plot=TRUE, boxes=FALSE) #l1 = arrows for environment
#par(new=T)
s.label(rlq1$lQ, boxes = FALSE) #ordination of the species distances, drawn without boxes around the names

plot(fourth, x.rlq=rlq1, alpha = 0.05, stat = "D2", type = "biplot")

```

Additional analyses also from esapubs
```{r}
nrepet=9999
testrlq<-randtest(rlq1, modeltype=6, nrepet = nrepet) #test based on total inertia of RLQ analysis to evaluate global significance of traits-environment relationship
testrlq
plot(testrlq)
Srlq<-fourthcorner2(env, species, traits, modeltype=6, p.adjust.method.G="fdr", nrepet=nrepet)
Srlq$trRLQ #returns the total inertia of the RLQ analysis


```
#this should be repeated after pruning the environmental and trait variables somewhat.  too many; after correcting for multiple comparisons everything washes out - wait until I can add my own pH and sand % measures.


Sample code from the package description
coa1<- dudi.coa(on a site x species abundance matrix with species in rows)
dudimil<- dudi.hillsmith(on a site x environment matrix with environment in rows)
duditrait<- dudi.hillsmith(on a species x trait matrix with traits in rows)
rlq1 <-rlq (dudimil, coa1, duditrait)

```{r}
  data(aviurba)
   coa1 <- dudi.coa(aviurba$fau, scannf = FALSE, nf = 2)
   dudimil <- dudi.hillsmith(aviurba$mil, scannf = FALSE, nf = 2, row.w = coa1$lw)
   duditrait <- dudi.hillsmith(aviurba$traits, scannf = FALSE, nf = 2, row.w = coa1$cw)
   rlq1 <- rlq(dudimil, coa1, duditrait, scannf = FALSE, nf = 2)
   plot(rlq1)
   summary(rlq1)
   randtest(rlq1)
   fourthcorner.rlq(rlq1,type="Q.axes")
   fourthcorner.rlq(rlq1,type="R.axes")

```


